Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-05-10 manuscript-figures.Rmd
Update manuscript-figures.Rmd

# Helper function: add a "point at infinity" for plotting when sigmaH is infinite
inf_skew_func <- function(df, n_rows_to_add) {
  # Value of sigmaH to replace 'Inf' with for plotting
  infinite_skew <- 10^3 + 2500

  # Get the largest value of log10(sigmaH)
  max_log10sigmaH <- filter(df, is.finite(sigmaH)) %>%
    select(sigmaH) %>%
    max() %>%
    log10()

  # add a second proxy sigmaH point at infinity to make the values easier to see there
  temp <- filter(df, is.infinite(sigmaH)) %>%
    mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
  # Create a proxy sigmaH variable that replaces Inf with infinite_skew
  df <- mutate(df, sigmaH_proxy = ifelse(is.finite(sigmaH),
                                         sigmaH, infinite_skew
  ))

  for (i in 1:n_rows_to_add) {
    bump_val <- 1+0.05*(i-1)
    temp <- filter(df, is.infinite(sigmaH)) %>%
      mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                                   mean(c(bump_val * infinite_divide, infinite_skew))
      ))
    df <- add_row(df, temp) %>% distinct()
    }

@@

# Get the largest value of log10(sigmaH)
max_log10sigmaH <- filter(AllOutputs, is.finite(sigmaH)) %>%
  select(sigmaH) %>%
  max() %>%
  log10()

# add a second proxy sigmaH point at infinity to make the values easier to see there
temp <- filter(AllOutputs, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# Create a proxy sigmaH variable that replaces Inf with infinite_skew
AllOutputs <- mutate(AllOutputs, sigmaH_proxy = ifelse(is.finite(sigmaH),
                                                       sigmaH, infinite_skew
))
AllOutputs <- add_row(AllOutputs, temp) %>% distinct()

# Define sigmaH tick breaks and labels
sigmaH_breaks <- c(10^c(seq(0, 3, by = 1)), infinite_divide, infinite_skew)
sigmaH_labels <- c(paste(c(10^c(seq(0, 3, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))

## Parasite thermal characteristics
ThermPathChars <- read_csv("results/ThermalCharacteristics.csv", show_col_types = FALSE)

# Title names for each mosquito+pathogen combination
ThermPathChars <- ThermPathChars %>%
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti+DENV" ~ "(a) *Ae. aegypti* & DENV",
    system_ID == "Aedes aegypti+ZIKV" ~ "(b) *Ae. aegypti* & ZIKV",
    system_ID == "Aedes albopictus+DENV" ~ "(c) *Ae. albopictus* & DENV",
    system_ID == "Anopheles spp.+P.fal." ~ "(d) *Anopheles* spp. & *P. falciparum*",
    system_ID == "Culex quinquefasciatus+WNV" ~ "(e) *Cx. quinquefasciatus* & WNV"
  ))
## Add a "point at infinity" for sigmaH (NB: to be added to get-analysis-dfs.R)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^3 + 2500

# Get the largest value of log10(sigmaH)
max_log10sigmaH <- filter(ThermPathChars, is.finite(sigmaH)) %>%
  select(sigmaH) %>%
  max() %>%
  log10()

# Create a proxy sigmaH variable that replaces Inf with infinite_skew
ThermPathChars <- mutate(ThermPathChars, sigmaH_proxy = ifelse(is.finite(sigmaH),
                                                               sigmaH, infinite_skew
))
# add a second proxy sigmaH point at infinity to make the values easier to see there
temp <- filter(ThermPathChars, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))

ThermPathChars <- add_row(ThermPathChars, temp) %>%
  arrange(sigmaH_proxy) %>%
  distinct()

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
Toptalt_df <- inf_skew_func(Toptalt_df, 6)

# Define sigmaH tick breaks and labels
sigmaH_breaks <- c(10^c(seq(-1, 2, by = 1)), infinite_divide, 10^log10(mean(c(infinite_divide,infinite_skew))))
sigmaH_labels <- c(paste(c(10^c(seq(-1, 2, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))

@@

# Plot mean of Topt heat map
Toptalt_plots <- Toptalt_df2 %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = transmission thermal optimum
  ggplot(aes(x = KH, y = sigmaH_proxy, z = median)) +
  # Topt for FINITE sigmaH:
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy < infinite_divide**0.99),
    breaks = Topt_breaks
  ) +
  # Topt for INFINITE sigmaH:
  # take mean of mean to hide variation from artifacts of uncertainty 
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy > infinite_divide**1.01) %>% 
      group_by(system_ID) %>% 
      mutate(mean = mean(mean), .groups = "keep"),
    size = 1,
    breaks = Topt_breaks
  ) +
  # # R0 = 1 contour: cheat and use where CTwidth = 0 (i.e. where R0<1 at any temperature)
  geom_contour(data = filter(data.CT, variable == "CTwidth"),
               aes(y = sigmaH, z = mean, colour = "R0 = 1"), 
               breaks = min(filter(data.CT, variable == "CTwidth", mean > 0)$mean),
               size = 1) +
  # geom_contour(data = data.R0_norm,
  #              aes(y = sigmaH, z = mean, colour = "R0=1"), breaks = c(1), size = 1) +
  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +

@@

```{r fig-6_CT_Width, dev = c('pdf','svg', 'tiff'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Width} The parasite population thermal tolerance range, the range of temperatures at which parasite transmission is expected to persist, is widest when biting tolerance is high and vertebrate host population density is low.'}
# Shows how the parasite thermal niche (temperatures between CT_min and CT_max) depends on vertebrate host availability (density and biting tolerance) in each of the systems we consider.

@@

CTWidth_df <- ThermPathChars %>%
  ungroup() %>%
  # Calculate thermal niche width
  mutate(CTwidth = CTmax - CTmin) %>%
  mutate(CTwidth = replace(CTwidth, is.na(CTwidth), 0)) %>%
  distinct() %>%
  select(system_label, sigmaH, sigmaH_proxy, KH, CTwidth) %>%
  # Arrange along the plotting variables
  arrange(system_label, sigmaH_proxy, KH, CTwidth)

# add many third proxy sigmaH points at infinity to make the values easier to see there
temp <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.25 * infinite_divide, infinite_skew))
  ))
temp2 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.15 * infinite_divide, infinite_skew))
  ))
temp3 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.05 * infinite_divide, infinite_skew))
  ))
temp4 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.3 * infinite_divide, infinite_skew))
  ))
temp5 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.2 * infinite_divide, infinite_skew))
  ))
temp6 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(1.1 * infinite_divide, infinite_skew))
  ))
temp7 <- filter(CTWidth_df, is.infinite(sigmaH)) %>%
  mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                               mean(c(infinite_divide, infinite_skew))
  ))

CTWidth_df <- add_row(CTWidth_df, temp)
CTWidth_df <- add_row(CTWidth_df, temp2)
CTWidth_df <- add_row(CTWidth_df, temp3)
CTWidth_df <- add_row(CTWidth_df, temp4)
CTWidth_df <- add_row(CTWidth_df, temp5)
CTWidth_df <- add_row(CTWidth_df, temp6)
CTWidth_df <- add_row(CTWidth_df, temp7) %>%
  arrange(sigmaH_proxy) %>%
  distinct()

CTWidth_heat_plots <- CTWidth_df %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = width of parasite thermal niche
  ggplot(aes(x = KH, y = sigmaH_proxy)) +

  # CTwidth for FINITE sigmaH:
  geom_contour_filled(
    data = filter(CTWidth_df, sigmaH_proxy < infinite_divide**0.98),
    aes(z = CTwidth)
  ) +

  # CTwidth for INFINITE sigmaH:
  geom_contour_filled(
    data = filter(CTWidth_df, sigmaH_proxy > infinite_divide**1.02),
    aes(z = CTwidth)
  ) +

  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +
  # color:
  scale_fill_viridis_d(
    name = "Parasite thermal tolerance\nrange width (°C)",
    option = "plasma"
  ) +

  # legend:
  guides(fill = guide_coloursteps(
    title.position = "top", title.hjust = 0.5,
    barwidth = unit(5, "cm"),
    show.limits = TRUE
  )) +

  # theme options:
  theme_minimal(11) +
  theme(
    legend.box = "horizontal",
    legend.position = "bottom",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.direction = "horizontal",
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

# Plot
shift_legend(CTWidth_heat_plots)
```

>>>>

# Helper function: add a "point at infinity" for plotting when sigmaH is infinite
inf_skew_func <- function(df, skew, divide, n_rows_to_add) {
  # Get the largest value of log10(sigmaH)
  max_log10sigmaH <- filter(df, is.finite(sigmaH)) %>%
    select(sigmaH) %>%
    max() %>%
    log10()

  # Used to create a gap dividing finite and infinite sigmaH
  infinite_divide <- 10^mean(log10(c(divide, skew)))

  # Create a proxy sigmaH variable that replaces Inf with skew
  df <- mutate(df, sigmaH_proxy = ifelse(is.finite(sigmaH),
                                         sigmaH, skew
  ))

  added_sigma_vals <- seq(divide, skew, length.out = n_rows_to_add)

  for (i in 1:n_rows_to_add) {
    bump_val <- 1+0.05*(i-1)
    temp <- filter(df, is.infinite(sigmaH)) %>%
      mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH,
                                   added_sigma_vals[i]))

    df <- add_row(df, temp) %>% distinct()
    }

@@

# ## Add a "point at infinity" for sigmaH
# # Value of sigmaH to replace 'Inf' with for plotting
# infinite_skew <- 10^3 + 2500
# 
# # Get the largest value of log10(sigmaH)
# max_log10sigmaH <- filter(AllOutputs, is.finite(sigmaH)) %>%
#   select(sigmaH) %>%
#   max() %>%
#   log10()
# 
# # Used to create a gap dividing finite and infinite sigmaH
# infinite_divide <- 10^mean(c(3, log10(infinite_skew)))
# 
# # add a second proxy sigmaH point at infinity to make the values easier to see there
# temp <- filter(AllOutputs, is.infinite(sigmaH)) %>%
#   mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# # Create a proxy sigmaH variable that replaces Inf with infinite_skew
# AllOutputs <- mutate(AllOutputs, sigmaH_proxy = ifelse(is.finite(sigmaH),
#                                                        sigmaH, infinite_skew
# ))
# AllOutputs <- add_row(AllOutputs, temp) %>% distinct()
# 
# # Define sigmaH tick breaks and labels
# sigmaH_breaks <- c(10^c(seq(0, 3, by = 1)), infinite_divide, infinite_skew)
# sigmaH_labels <- c(paste(c(10^c(seq(0, 3, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))
# 
# ## Parasite thermal characteristics
# ThermPathChars <- read_csv("results/ThermalCharacteristics.csv", show_col_types = FALSE)
# 
# # Title names for each mosquito+pathogen combination
# ThermPathChars <- ThermPathChars %>%
#   mutate(system_label = case_when(
#     system_ID == "Aedes aegypti+DENV" ~ "(a) *Ae. aegypti* & DENV",
#     system_ID == "Aedes aegypti+ZIKV" ~ "(b) *Ae. aegypti* & ZIKV",
#     system_ID == "Aedes albopictus+DENV" ~ "(c) *Ae. albopictus* & DENV",
#     system_ID == "Anopheles spp.+P.fal." ~ "(d) *Anopheles* spp. & *P. falciparum*",
#     system_ID == "Culex quinquefasciatus+WNV" ~ "(e) *Cx. quinquefasciatus* & WNV"
#   ))
# ## Add a "point at infinity" for sigmaH (NB: to be added to get-analysis-dfs.R)
# # Value of sigmaH to replace 'Inf' with for plotting
# infinite_skew <- 10^3 + 2500
# 
# # Get the largest value of log10(sigmaH)
# max_log10sigmaH <- filter(ThermPathChars, is.finite(sigmaH)) %>%
#   select(sigmaH) %>%
#   max() %>%
#   log10()
# 
# # Used to create a gap dividing finite and infinite sigmaH
# infinite_divide <- 10^mean(c(3, log10(infinite_skew)))
# 
# # Create a proxy sigmaH variable that replaces Inf with infinite_skew
# ThermPathChars <- mutate(ThermPathChars, sigmaH_proxy = ifelse(is.finite(sigmaH),
#                                                                sigmaH, infinite_skew
# ))
# # add a second proxy sigmaH point at infinity to make the values easier to see there
# temp <- filter(ThermPathChars, is.infinite(sigmaH)) %>%
#   mutate(sigmaH_proxy = ifelse(is.finite(sigmaH), sigmaH, 1.2 * infinite_divide))
# 
# ThermPathChars <- add_row(ThermPathChars, temp) %>%
#   arrange(sigmaH_proxy) %>%
#   distinct()
# 
# ## Options
# # Define sigmaH tick breaks and labels
# sigmaH_breaks <- c(10^c(seq(-1, 3, by = 1)), infinite_divide, infinite_skew)
# sigmaH_labels <- c(paste(c(10^c(seq(-1, 3, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))

@@

# add many proxy sigmaH points at infinity to make the values easier to see there
max_sigmaH <- max(filter(Toptalt_df, is.finite(sigmaH))$sigmaH)
# Value of sigmaH to replace 'Inf' with for plotting
infinite_skew <- 10^log10(max_sigmaH*1.5) # 1.1*max_sigmaH #10^2 + 80
# Used to create a gap dividing finite and infinite sigmaH
infinite_divide <- 10^log10(max_sigmaH*1)

# infinite_skew <- 10^2 + 150
Toptalt_df <- inf_skew_func(Toptalt_df, infinite_skew, infinite_divide, 8)

# infinite_divide <- 10^mean(c(2, log10(infinite_skew)))

# Define sigmaH tick breaks and labels
sigmaH_breaks <- c(10^c(seq(-1, 2, by = 1)), mean(c(infinite_divide, infinite_skew)), max(Toptalt_df$sigmaH_proxy))
sigmaH_labels <- c(paste(c(10^c(seq(-1, 2, by = 1)))), TeX("$\\ldots$"), TeX("$\\infty$"))

@@

# Plot mean of Topt heat map
Toptalt_plots <- Toptalt_df %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = transmission thermal optimum
  ggplot(aes(x = KH, y = sigmaH_proxy, z = median)) +
  # Topt for FINITE sigmaH:
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy < infinite_divide**0.99),
    breaks = Topt_breaks
  ) +
  # Topt for INFINITE sigmaH:
  # take mean of mean to hide variation from artifacts of uncertainty 
  geom_contour_filled(
    data = filter(Toptalt_df, sigmaH_proxy > infinite_divide**1.02) %>% 
      group_by(system_ID) %>% 
      mutate(mean = mean(mean), .groups = "keep"),
    size = 1,
    breaks = Topt_breaks
  ) +
  # # R0 = 1 contour: cheat and use where CTwidth = 0 (i.e. where R0<1 at any temperature)
  geom_contour(data = filter(data.CT, variable == "CTwidth"),
               aes(y = sigmaH, z = mean, colour = "R0 = 1"), 
               breaks = min(filter(data.CT, variable == "CTwidth", mean > 0)$mean),
               size = 1) +
  # geom_contour(data = data.R0_norm,
  #              aes(y = sigmaH, z = mean, colour = "R0=1"), breaks = c(1), size = 1) +
  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +

@@

```{r fig-6_CT_Width, dev = c('pdf','svg', 'tiff'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Width} The parasite population thermal tolerance range, the range of temperatures at which parasite transmission is expected to persist, is widest when biting tolerance is high and vertebrate host population density is low.'}
# Shows how the parasite thermal niche (temperatures between CT_min and CT_max) 
# depends on vertebrate host availability (density and biting tolerance) in each
# of the focal systems.

@@

CTwidth_df <- data.CT %>%
  filter(variable == "CTwidth") %>% 
  mutate(sigmaH_proxy = sigmaH) %>% 
  # Arrange along the plotting variables
  arrange(system_ID, KH, sigmaH_proxy, mean) %>% 
  ungroup()

infinite_skew <- 10^3 + 250
CTwidth_df <- inf_skew_func(CTwidth_df, infinite_skew, divide_order = 3, 8)

infinite_divide <- 10^mean(c(2, log10(infinite_skew)))
CTWidth_heat_plots <- CTwidth_df %>%
  ## Set up plot ##
  # x = biting tolerance (with a point at infinity),
  # y = vertebrate host population density,
  # color = width of parasite thermal mean_CTwidth
  ggplot(aes(x = KH, y = sigmaH_proxy, z = mean)) +
  # CTwidth for FINITE sigmaH:
  geom_contour_filled(data = filter(CTwidth_df, sigmaH_proxy < infinite_divide**0.99)) +
  # CTwidth for INFINITE sigmaH:
  geom_contour_filled(data = filter(CTwidth_df, sigmaH_proxy > infinite_divide**1.01)) +
  # x-axis: log10-scale with no buffer space
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E4),
    breaks = 10^seq(-1, 4, 1),
    labels = TeX(c("$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  # y-axis: log10-scale y axis, with no buffer space
  scale_y_log10(
    name = TeX("Biting tolerance (bites per host per day)"),
    expand = c(0, 0),
    breaks = sigmaH_breaks,
    labels = sigmaH_labels
  ) +
  # color:
  scale_fill_viridis_d(
    name = "Mean parasite\nthermal tolerance\nrange width (°C)",
    option = "plasma"
  ) +
  # faceting:
  facet_wrap(~system_ID, scales = "free", nrow = 2, labeller = labeller()) +
  # legend:
  guides(fill = guide_coloursteps(
    title.position = "top", title.hjust = 0.5,
    barwidth = unit(5, "cm"),
    show.limits = TRUE
  )) +
  # theme options:
  theme_minimal(16) +
  theme(
    legend.box = "horizontal",
    legend.position = "bottom",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.direction = "horizontal",
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

# Plot
CTWidth_heat_plots <- shift_legend(CTWidth_heat_plots)
